---
title: "Kaggle Titanic"
output: html_notebook
---
This is a replica of [Megan Risdal's post, Exploring Survival on the Titanic
](https://www.kaggle.com/mrisdal/exploring-survival-on-the-titanic), with a bit more [tidyverse](https://www.tidyverse.org/) style to make the code clearer.

# Load Packages and Data
```{r message=FALSE, warning=FALSE}
# Load packages
library(tidyverse)
library(mice)
library(randomForest)
library(stringr)
library(scales)

# Load data
train <- read_csv('./input/train.csv');
test <- read_csv('./input/test.csv');
full <- bind_rows(train=train, test=test, .id = 'dataset');
```

# Feature Engineering
## People's Titles
#### Extract title from names
```{r}
full <- full %>% mutate(Title=str_replace_all(Name, '(.*, )|(\\..*)', ''))
table(full$Sex, full$Title)
```

#### Duplicated titles
Some titles mean the same thing, like Miss = Ms = Mlle, and Mme = Mrs. Combine them.
```{r}
full <- full %>% 
    mutate_at(.vars = vars(Title), 
              .funs = funs(ifelse(. %in% c('Mlle','Ms'), 'Miss', .))) %>% 
    mutate_at(.vars = vars(Title),
              .funs = funs(ifelse(. %in% c('Mme'), 'Mrs', .)))
```

#### Rare titles
Some titles are quite rare. Combine them to one label, 'Rare Title'
```{r}
title_counts <- table(full$Title)
rare_title <- names(title_counts)[title_counts<10]
full <- full %>% mutate_at(.vars = vars(Title),
                           .funs = funs(ifelse(. %in% rare_title, 'Rare Title', .)))
table(full$Sex, full$Title)
```


## Families
#### Family names
Get family names from full names
```{r}
full$Surname <- full$Name %>% str_split('[,.]') %>% lapply('[[', 1) %>% unlist() %>% trimws()
```


#### Family size
Get family sizes from counts of sibling, spouse, parents, and children
```{r}
full <- full %>% 
    mutate(Fsize = SibSp + Parch +1) %>% 
    mutate(Family = paste(Surname, Fsize, sep='_'))
ggplot(data = subset(full, !is.na(Survived)),
       aes(x = Fsize, fill=factor(Survived))) +
    geom_bar(stat = 'count', position = 'dodge') +
    scale_x_continuous(breaks = 1:11) +
    labs(x = 'Family Size') 
```

#### Singleton, samll, and large families
Categorize family sizes into 'singleton', 'small', and 'large'
```{r}
full <- full %>% mutate(FsizeD = case_when(
    Fsize==1 ~ 'singleton',
    Fsize>4 ~ 'large',
    TRUE ~ 'small'
))
mosaicplot(table(full$FsizeD, full$Survived), main='Family Size vs Survival', shade = TRUE)

```
Looks like small families are more likely to survive.

## Deck
Deck information is available from the `Cabin` code
```{r}
full$Deck <-  full$Cabin %>% substr(1,1) %>% as.factor()
```

## Missing Values
#### Count `NA`'s
```{r}
full %>% is.na() %>% colSums()
```

#### Missing `Embarked`
```{r}
selector <- full$Embarked %>% is.na()
full[selector, ]
```
We can see that these two passengers with missing `Embarked` variables were in 1st class (`Pclass=1`), and both paid $80.
Lets' take a look at what fare passengers from different embark point and different class paid.
```{r message=FALSE, warning=FALSE}
ggplot(data = subset(full, !selector),
       aes(x=Embarked, y=Fare, fill=factor(Pclass))) +
    geom_boxplot() +
    scale_y_continuous()
full %>% filter(!selector, Embarked=='C', Pclass==1) %>% select(Fare) %>% apply(2, median)
```

We can see that the fare these passengers paid ($80) is close to the median of what 1st class passengers from C paid, so we infer that they embarded from C (Cherbourg)
```{r}
full$Embarked[selector] <- 'C'
full %>% is.na() %>% colSums()
```

#### Missing `Fare`
There is a passenger with missing Fare
```{r}
selector <- full$Fare %>% is.na()
full[selector,]
```
This is a 3rd class passenger embarked form `S`. The `Fare` distribution of this group of passengers is
```{r}
group_selector <- (!selector) & (full$Embarked=='S') & (full$Pclass==3)
ggplot(data = filter(full, group_selector),
       aes(x=Fare)) +
    geom_density(fill='deepskyblue')
```

Use median fare of this class group to fill the missing fare
```{r}
full$Fare[selector] <- full$Fare[group_selector] %>% median()
full %>% is.na() %>% colSums()

```

#### Missing `Age`
Use the ['mice' package](https://cran.r-project.org/web/packages/mice/mice.pdf) to impute `Age`. MICE stands for Multivariate Imputation by Chained Equations.
```{r include=FALSE}
set.seed(129)
unuseful_features <- c('PassengerId', 'Name', 'Ticket', 'Cabin', 'Family', 'Surname', 'Survived')
mice_mod <- mice(full[, !names(full) %in% unuseful_features], method = 'rf')
mice_output <- mice_mod %>% complete()
```

Plot age distributions, to see if `mice_output` appears normal
```{r}
par(mfrow=c(1,2))
hist(full$Age, freq = F, main='Age: Original Data', col = 'darkgreen', ylim = c(0, 0.04))
hist(mice_output$Age, freq = F, main = 'Age: MICE Output', col = 'lightgreen', ylim = c(0, 0.04))
```
Use `mice_output$Age` to replace `Age` in the original dataset.
```{r}
full$Age <- mice_output$Age
full$Age %>% is.na() %>% sum()
```

## More new features
#### Adult/child
```{r}
full <- full %>% mutate(Child=ifelse(Age<18, 'Child', 'Adult') %>% as.factor())
table(full$Child, full$Survived)
mosaicplot(.Last.value, shade=TRUE, main='Child/Adult vs. Survival')
```

#### Mother
```{r}
full <- full %>% mutate(Mother=(Sex=='female' & Parch>0 & Age>18 & Title!='Miss'))
table(full$Mother, full$Survived)
mosaicplot(.Last.value, shade=TRUE, main='Mother (T/F) vs. Survival')
```


Now we don't have any missing values except for `Cabin` and `Deck`
```{r}
full %>% is.na() %>% colSums()
```

## Final Cleaning: Strings to Factors
```{r}
cols <- c('Pclass', 'Sex', 'Embarked', 'Title', 'FsizeD', 'Child', 'Mother')
full[cols] <- full[cols] %>% lapply(factor)
sapply(full, class)
```


# Model
## Train-test split
Split the full dataset back into `train` and `test` sets
```{r}
train <- full %>% filter(dataset=='train') %>% select(-dataset)
test  <- full %>% filter(dataset=='test')  %>% select(-dataset)
```


## Random Forrest model
```{r}
set.seed(754)
rf_model <- randomForest(factor(Survived) ~ Pclass + Sex + Age + SibSp + 
                                            Parch + Fare + Embarked + Title + 
                                            FsizeD + Child + Mother,
                         data = train)
par(mfrow=c(1,1))
plot(rf_model, ylim=c(0, 0.36))
legend('topright', colnames(rf_model$err.rate), col = 1:3, fill = 1:3)
```


## Variable Importance
```{r}
importance <- importance(rf_model)
varImportance <- data.frame(Variables = row.names(importance),
                            Importance = round(importance[,'MeanDecreaseGini'], 2))
rankImportance <- varImportance %>% 
    mutate(Rank = dense_rank(desc(Importance))) %>% 
    arrange(Rank)
ggplot(data = rankImportance,
       aes(x = reorder(Variables, Importance), y = Importance, fill=Importance)) +
    geom_bar(stat = 'identity') +
    geom_text(aes(x=Variables, y=0.5, label=paste0('#',Rank)), hjust=-1, vjust=0.55, color='red') +
    coord_flip()
```




# Prediction
```{r}
pred <- predict(rf_model, test)
solution <- tibble(PassengerId = test$PassengerId, Survived = pred)
write_csv(solution, 'rf_mod_Solution.csv')
```



# Summary
## Workflow Overview
- Check data
    - Shape
    - Feature types (categorical/continuous)
    - Missing values
- Data cleaning + feature engineering
- Modeling
- Prediction

## What can be learned
- Feature engineering. E.g. extracting titles from names
- Imputation of missing data: `mice`
- Categorical features visualization with `mosaicplot()`

## What can be improved
- Model evaluation
- Random forest parameter tuning




